### 2022-11-22 paper-thesis/chapters/20-adding-booleans.tex
Finish one chapter

Next we will add \verb|Postulate| to the language. A postulate declares that there is a variable with a certain type, without specifying a value. Through the view of the Curry-Howard correspondence, this is equivalent to an axiom. At the current stage, this is useful for testing the language. For example, this is a test with a postulate:

\begin{lstlisting}
postulate T : Type;
if true then Bool else T end
\end{lstlisting}
The following rules are used to implement the feature, which translate cleanly to Statix:

@@

Finally, we will add \verb|TypeAssert| to the language.

@@

>>>>

Next we will add \verb|Postulate| to the language. A postulate declares that there is a variable with a certain type, without specifying a value. Through the view of the Curry-Howard correspondence, this is equivalent to an axiom. At the current stage, this is useful for testing the language. For example, this is a test with a postulate:
\begin{lstlisting}
postulate T : Type;
if true then Bool else T end
\end{lstlisting}
The following rules are used to implement the feature, which translate cleanly to Statix:

@@

Finally, we will add \verb|TypeAssert| to the language. This is another feature that is useful for testing. It takes an expression, and it asserts that the expression has a certain type. For example, we can do the following:
\begin{lstlisting}
postulate T : Type;
true : if true then Bool else T end
\end{lstlisting}

Implementing this feature is also straight-forward, we add the following two rules, which translate cleanly to Statix:

\begin{figure}[ht]
	\begin{mathpar}
		\inferrule{
			\bhr
			{\scope{s}{b}}
			{a}
			{\scope{s'}{b'}}
		}{
			\bhr
			{\scope{s}{\co{TypeAssert}(b, t)}}
			{a}
			{\scope{s'}{b'}}
		}

		\inferrule{
			\toe{s}{t}{tt}
			\\\beq{tt}{\co{Type}()}
			\\\\\toe{s}{b}{bt}
			\\\beq
			{bt}
			{\scope{s}{t}}
		} {
			\toe{s}{\co{TypeAssert}(b, t)}{bt}
		}
	\end{mathpar}
	\caption{Rules for type assertions}
	\label{fig:typeassert-rules}
\end{figure}

@@

Now that a few features have been implemented, we can discuss how easy the language is to extend. The conclusion so far is that it is doable in a clean way. Specifically, the approach for each of the features above was:

\begin{enumerate}
	\item Define the parsing rules for the new feature
	\item Create a new file, \verb|tp_bools.stx| and import this file in the main \verb|type_check.stx| file.
	\item In this file, define a case for the \verb|betaReduceHead| rule for the constructors that were added. Also define cases for \verb|expectBetaEq| and \verb|betaReduce| if necessary.
	\item In this file, define a case for the \verb|typeOfExpr| rule for the constructors that were added.
\end{enumerate}

This allows each feature to be isolated to its own file. If we decide that we don't like a feature after all, we can remove it simply by unimporting the file. This problem is similar to the expression problem\cite{expression_problem}, which Statix solves very well.

In the following chapters, we're going to be extending the language further with these more interesting features using this approach:
\begin{itemize}
	\item Inference (chapter \ref{ch:inference})
	\item Inductive Datatypes (chapter \ref{ch:datatypes})
	\item Universes (chapter \ref{ch:universes})
\end{itemize}
